---
title: "Introduction to Model Deployments"
author: "Peter Hurford"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes  
vignette: >
  %\VignetteIndexEntry{Using Model Deployments}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

In DataRobot, a **model deployment** is the act of launching a DataRobot model to a dedicated prediction cluster using the model monitoring system. It is still possible to make predictions with models that have not been deployed in this way, but users will not be able to access the model monitoring information available via this API without deploying via the model monitoring system.


## Retrieving Model Deployments

Before you can retrieve information on a model deployment, you must first request the deployment. This requires deploying a specific model from a specific DataRobot project, and giving the deployment an informative label to describe the model.

```{r results = "asis", message = FALSE, warning = FALSE, eval = FALSE}
projectId <- GetProjectList()$projectId[[1]]
model <- ListModels(projectId)[[1]]
RequestModelDeployment(projectId, model$modelId, label = "GBM Model for Loan Decisioning")
```

You can then get information on all model deployments using `ListModelDeployments()`. Note that this function lists information on all deployments across all projects.

```{r results = "asis", message = FALSE, warning = FALSE, eval = FALSE}
print(ListModelDeployments()[[1]])
```

```{r results = "asis", echo = FALSE}
modelDeployments <- readRDS("listModelDeployments.rds")
print(modelDeployments[[1]])
```

```{r results = "asis", message = FALSE, warning = FALSE, eval = FALSE}
kable(head(summary(ListModelDeployments())))
```

```{r results = "asis", echo = FALSE}
library(knitr)
modelDeployments <- readRDS("listModelDeployments.rds")
kable(head(summary(modelDeployments)))
```

You can use the `limit` param to only return a certain amount of model deployments.

```{r results = "asis", message = FALSE, warning = FALSE, eval = FALSE}
modelDeployments <- ListModelDeployments(limit = 2)
kable(head(summary(modelDeployments)), longtable = TRUE, booktabs = TRUE, row.names = TRUE)
```

```{r results = "asis", echo = FALSE}
modelDeployments <- readRDS("listModelDeploymentsLimit2.rds")
kable(head(summary(modelDeployments)), longtable = TRUE, booktabs = TRUE, row.names = TRUE)
```

You can use `orderBy` to determine the order that model deployments are retrieved (and limited).

```{r results = "asis", message = FALSE, warning = FALSE, eval = FALSE}
modelDeployments <- ListModelDeployments(limit = 2, orderBy = "-label")
kable(head(summary(modelDeployments)), longtable = TRUE, booktabs = TRUE, row.names = TRUE)
```

```{r results = "asis", echo = FALSE}
modelDeployments <- readRDS("listModelDeploymentsLimit2OrderByLabel.rds")
kable(head(summary(modelDeployments)), longtable = TRUE, booktabs = TRUE, row.names = TRUE)
```

You can use `offset` to skip that many deployments.

```{r results = "asis", message = FALSE, warning = FALSE, eval = FALSE}
modelDeployments <- ListModelDeployments(offset = 2)
kable(head(summary(modelDeployments)), longtable = TRUE, booktabs = TRUE, row.names = TRUE)
```

```{r results = "asis", echo = FALSE}
modelDeployments <- readRDS("listModelDeploymentsOffset2.rds")
kable(head(summary(modelDeployments)), longtable = TRUE, booktabs = TRUE, row.names = TRUE)
```

You can use `query` to search for a particular model deployment via searching the label and description.

```{r results = "asis", message = FALSE, warning = FALSE, eval = FALSE}
modelDeployments <- ListModelDeployments(query = "Decisioning")
kable(head(summary(modelDeployments)), longtable = TRUE, booktabs = TRUE, row.names = TRUE)
```

```{r results = "asis", echo = FALSE}
modelDeployments <- readRDS("listModelDeploymentsQuery.rds")
kable(head(summary(modelDeployments)), longtable = TRUE, booktabs = TRUE, row.names = TRUE)
```

You can use `status` to filter models to a particular model deployment status.

```{r results = "asis", message = FALSE, warning = FALSE, eval = FALSE}
modelDeployments <- ListModelDeployments(status = "Active")
kable(head(summary(modelDeployments)), longtable = TRUE, booktabs = TRUE, row.names = TRUE)
```

```{r results = "asis", echo = FALSE}
modelDeployments <- readRDS("listModelDeployments.rds")
kable(head(summary(modelDeployments)), longtable = TRUE, booktabs = TRUE, row.names = TRUE)
```

Lastly, if you have a specific model deployment ID, you can use `GetModelDeployment` to fetch information on that particular deployment.

```{r results = "asis", message = FALSE, warning = FALSE, eval = FALSE}
modelDeployments <- ListModelDeployments()
modelDeployment <- GetModelDeployment(modelDeployments[[1]]$id)
print(modelDeployment)
```

```{r results = "asis", echo = FALSE}
modelDeployment <- readRDS("getModelDeployment.rds")
print(modelDeployment)
```

## Updating a Model Deployment

After a model deployment is made, the `label`, `description`, and `status` can be updated using `UpdateModelDeployment`. Here, the label and the status are both changed:

```{r results = "asis", message = FALSE, warning = FALSE, eval = FALSE}
modelDeployments <- ListModelDeployments()
UpdateModelDeployment(modelDeployments[[1]]$id, label = "Archived GBM", status = "archived")
```

```{r results = "asis", echo = FALSE}
modelDeployment <- readRDS("updatedModelDeployment.rds")
print(modelDeployment)
```


## Model Deployment Service Statistics

```{r results = "asis", message = FALSE, warning = FALSE, eval = FALSE}
modelDeployments <- ListModelDeployments()
serverHealth <- GetModelDeploymentServiceStatistics(modelDeployments[[1]]$id)
print(serverHealth)
```

```{r results = "asis", echo = FALSE}
serverHealth <- readRDS("serverHealth.rds")
print(serverHealth)
```


## Model Deployment Action Log

```{r results = "asis", message = FALSE, warning = FALSE, eval = FALSE}
modelDeployments <- ListModelDeployments()
actionLog <- GetModelDeploymentActionLog(modelDeployments[[1]]$id)
print(actionLog)
```

```{r results = "asis", echo = FALSE}
actionLog <- readRDS("actionLog.rds")
print(actionLog)
```
